---
title: "Data Wrangling Task"
author: "Bradley Hughes"
date: "4/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rio)
library(tidyverse)
library(magrittr)
library(psych)
```
#Wrangling Hackathon
In this hackathon you will practice data wrangling by merging data from three csv files. One of these files is data collected in the lab after a dyadic interaction and the items of interest are BFI personality perceptions of the participant's interaction partner. All of the complete cases we want in our final data frame are in this data set. We will be connecting the self-report data collected prior to the lab study from two seperate dataframes in slightly different formats. You will be using different identifiers to connect each of these self-report to the lab data. At the end the dataframe will only have calculated BFI self-report and perception scores for the big five and HP. From there we will manipulate the dataframe so that it can be analyzed as dyads in lavaan, and then make it into tidy data(long).

##Calculating BFI-2 scores
As to maximize you time cleaning and wrangling the data, I have provided code that will calculate scores for each trait and save them to a new column in the dataframe. I did this with the scoreItems function in the psych package. It provides some reliability statitistics. 

##Data
In order to maximize the reproducability of the eventual analysis, begin by importing the raw data files directly from Qualtrics. All identifiable data has been removed or altered and all of the data has been simulated. However, besides a few missing columns it is exactly what qualtrics exports.

#Cleaning and Merging
##Start with the lab data. When this data frame is cleaned it will contain all the usable dyads in so much as we have the in-lab data for them.

```{r import and clean lab data}
#import data

#remove pilot data, Dyads 1-14


# exclude participants who knew each other before participating
#self-identified - already_know = 1 (they know the other participant) or well_known 1 or 2 (they have only seen them or interacted with them briefly)

#My RAs also identified dyads that were friends although they did not admit so in the survey. Remove dyads 17 and 25.

#exclude 2 participants for bringing emotional support animals to the study, dyads 110 and 124

#fixing some data entry errors:
#participant 45 in dyad 45 should be participant 23.

#participant 58 entered their last 4 wrong, should be 4058


#remove 2 columns of "know each other" variables

#identify any dyads with only one person and remove them

#print the head of the dataframe
```

##Next calculate BFI trait scores. For ease I have used the generic data in all of these functions, so please assign your dataframe to data below.

```{r score BFI-2 and subset dataframe}

data <-  

mitems <- cbind(data$BFII_2, data$BFII_7, data$BFII_12, data$BFII_17,
                       data$BFII_22, data$BFII_27, data$BFII_32, 
                        data$BFII_37, data$BFII_42, data$BFII_47, 
                        data$BFII_52, data$BFII_57)
mkey <- c(1 , 1, -1, -1, -1, 1, 1, -1, -1, -1, 1, 1) 
agree_per <- scoreItems(mkey, mitems, totals = FALSE, ilabels = NULL,missing=TRUE, impute="none", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$agree_per <- agree_per$scores

agree_per
sd(data$agree_per)

#Extraversion
# Ext = 1, 6, 11R, 16R, 21, 26R, 31R, 36R, 41, 46, 51R, 56
mitemse <- cbind(data$BFII_1, data$BFII_6, data$BFII_11, data$BFII_16,
                 data$BFII_21, data$BFII_26, data$BFII_31, data$BFII_36, 
                 data$BFII_41, data$BFII_46, data$BFII_51, data$BFII_56)
mkeye <- c(1 , 1, -1, -1, 1, -1, -1, -1, 1, 1, -1, 1)
extra_per <- scoreItems(mkeye, mitemse, totals = FALSE, ilabels = NULL,missing=TRUE, impute="mean", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$extra_per <- extra_per$scores

extra_per
sd(data$extra_per)

#openness
# Open = 5R, 10, 15, 20, 25R, 30R, 35, 40, 45R, 50R, 55R, 60
mitemso <- cbind(data$BFII_5, data$BFII_10, data$BFII_15, data$BFII_20,
                 data$BFII_25,data$BFII_30, data$BFII_35, data$BFII_40,
                 data$BFII_45, data$BFII_50, data$BFII_55, 
                 data$BFII_60)
mkeyo <- c(-1 , 1, 1, 1, -1, -1, 1, 1, -1, -1, -1, 1)
open_per <- scoreItems(mkeyo, mitemso, totals = FALSE, ilabels = NULL,missing=TRUE, impute="mean", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$open_per <- open_per$scores

open_per
sd(data$open_per)

#Conscientiousness
# Consc = 3R, 8R, 13, 18, 23R, 28R, 33, 38, 43, 48R, 53, 58R
mitemsc <- cbind(data$BFII_3, data$BFII_8, data$BFII_13, data$BFII_18, 
                 data$BFII_23, data$BFII_28, data$BFII_33, data$BFII_38, 
                 data$BFII_43, data$BFII_48, data$BFII_53, data$BFII_58)
mkeyc <- c(-1 , -1, 1, 1, -1, -1, 1, 1, 1, -1, 1, -1)
consc_per <- scoreItems(mkeyc, mitemsc, totals = FALSE, ilabels = NULL,missing=TRUE, impute="mean", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$consc_per <- consc_per$scores

consc_per
sd(data$consc_per)

#Neuroticism
# Neur = 4R, 9R, 14, 19, 24R, 29R, 34, 39, 44R, 49R, 54, 59
mitemsn <- cbind(data$BFII_4, data$BFII_9, data$BFII_14, data$BFII_19,
                 data$BFII_24, data$BFII_29, data$BFII_34, data$BFII_39, 
                 data$BFII_44, data$BFII_49, data$BFII_54, 
                 data$BFII_59)
mkeyn <- c(1 , -1, 1, 1, -1, -1, 1, 1, -1, -1, 1, 1)
neuro_per <- scoreItems(mkeyn, mitemsn, totals = FALSE, ilabels = NULL,missing=TRUE, impute="mean", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$neuro_per<- neuro_per$scores

neuro_per
sd(data$neuro_per)

# HP = HP45r HP46r HP47r HP48r HP49r HP50 HP51 HP52
mitemshp <- cbind(data$BFII_HP45, data$BFII_HP46, data$BFII_HP47,
                  data$BFII_HP48, data$BFII_HP49, data$BFII_HP50,
                  data$BFII_HP51, data$BFII_HP52)
mkeyhp <- c(-1 , -1, -1, -1, -1, 1, 1, 1)
hp_per <- scoreItems(mkeyhp, mitemshp, totals = FALSE, ilabels = NULL, missing=TRUE, impute="mean", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$hp_per <- hp_per$scores

hp_per
sd(data$hp_per)

#Now create a dataframe with only Dyad_ID, P_ID, ID, and the BFI trait scores.
##Be sure to check the class of the variables and make any adjustments necessary.
#We will use the P_ID and ID variables to match with the other dataframes. This makes it vital that they are of the same class, and in the same format.

#print the head of the dataframe
```

##Next, import and score data from online part 1 and prescreen. 
###Online

```{r online data}
#import online data

#score BFI

#Agreeableness
#alabel <- c(data$BFIS_2, data$BFIS_7, data$BFIS_12, data$BFIS_17,
 #                      data$BFIS_22, data$BFIS_27, data$BFIS_32, 
  #                      data$BFIS_37, data$BFIS_42, data$BFIS_47, 
   #                     data$BFIS_52, data$BFIS_57)
mitems <- cbind(data$BFIS_2, data$BFIS_7, data$BFIS_12, data$BFIS_17,
                       data$BFIS_22, data$BFIS_27, data$BFIS_32, 
                        data$BFIS_37, data$BFIS_42, data$BFIS_47, 
                        data$BFIS_52, data$BFIS_57)
mkey <- c(1 , 1, -1, -1, -1, 1, 1, -1, -1, -1, 1, 1)
agree_self <- scoreItems(mkey, mitems, totals = FALSE, ilabels = NULL,missing=TRUE, impute="mean", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$agree_self <- agree_self$scores

agree_self
sd(data$agree_self)


#Extraversion
# Ext = 1, 6, 11R, 16R, 21, 26R, 31R, 36R, 41, 46, 51R, 56
mitemse <- cbind(data$BFIS_1, data$BFIS_6, data$BFIS_11, data$BFIS_16,
                 data$BFIS_21, data$BFIS_26, data$BFIS_31, data$BFIS_36, 
                 data$BFIS_41, data$BFIS_46, data$BFIS_51, 
                 data$BFIS_56)
mkeye <- c(1 , 1, -1, -1, 1, -1, -1, -1, 1, 1, -1, 1)
extra_self <- scoreItems(mkeye, mitemse, totals = FALSE, ilabels = NULL,missing=TRUE, impute="mean", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$extra_self <- extra_self$scores

extra_self
sd(data$extra_self)

#openness
# Open = 5R, 10, 15, 20, 25R, 30R, 35, 40, 45R, 50R, 55R, 60
mitemso <- cbind(data$BFIS_5, data$BFIS_10, data$BFIS_15, data$BFIS_20,
                 data$BFIS_25, data$BFIS_30, data$BFIS_35, data$BFIS_40, 
                 data$BFIS_45, data$BFIS_50, data$BFIS_55, data$BFIS_60)
mkeyo <- c(-1 , 1, 1, 1, -1, -1, 1, 1, -1, -1, -1, 1)
open_self <- scoreItems(mkeyo, mitemso, totals = FALSE, ilabels = NULL,missing=TRUE, impute="mean", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$open_self <- open_self$scores

open_self
sd(data$open_self)

#Conscientiousness
# Consc = 3R, 8R, 13, 18, 23R, 28R, 33, 38, 43, 48R, 53, 58R
mitemsc <- cbind(data$BFIS_3, data$BFIS_8, data$BFIS_13, data$BFIS_18,
                 data$BFIS_23, data$BFIS_28, data$BFIS_33, data$BFIS_38, 
                 data$BFIS_43, data$BFIS_48, data$BFIS_53, 
                 data$BFIS_58)
mkeyc <- c(-1 , -1, 1, 1, -1, -1, 1, 1, 1, -1, 1, -1)
consc_self <- scoreItems(mkeyc, mitemsc, totals = FALSE, ilabels = NULL,missing=TRUE, impute="mean", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$consc_self <- consc_self$scores

consc_self
sd(data$consc_self)

#Neuroticism
# Neur = 4R, 9R, 14, 19, 24R, 29R, 34, 39, 44R, 49R, 54, 59
mitemsn <- cbind(data$BFIS_4, data$BFIS_9, data$BFIS_14, data$BFIS_19,
                 data$BFIS_24, data$BFIS_29, data$BFIS_34, data$BFIS_39, 
                 data$BFIS_44, data$BFIS_49, data$BFIS_54, data$BFIS_59)
mkeyn <- c(1 , -1, 1, 1, -1, -1, 1, 1, -1, -1, 1, 1)
neuro_self <- scoreItems(mkeyn, mitemsn, totals = FALSE, ilabels = NULL,missing=TRUE, impute="mean", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$neuro_self<- neuro_self$scores

neuro_self
sd(data$neuro_self)

# HP = HP45r HP46r HP47r HP48r HP49r HP50 HP51 HP52
mitemshp <- cbind(data$BFIS_HP45, data$BFIS_HP46, data$BFIS_HP47,
                  data$BFIS_HP48, data$BFIS_HP49, data$BFIS_HP50,
                  data$BFIS_HP51, data$BFIS_HP52)
mkeyhp <- c(-1 , -1, -1, -1, -1, 1, 1, 1)
hp_self <- scoreItems(mkeyhp, mitemshp, totals = FALSE, ilabels = NULL, missing=TRUE, impute="mean", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$hp_self <- hp_self$scores

hp_self
sd(data$hp_self)

#create a dataframe with only ID and scored self-report BFI, remove duplicate ID's. This means the participant took part 1 or began part 1 multiple times. 

#print the head of the dataframe
```

###Prescreen

```{r prescreen data}
#import prescreen data


#score prescreen data

#Agreeableness
#alabel <- c(data$BFIS_2, data$BFIS_7, data$BFIS_12, data$BFIS_17,
 #                      data$BFIS_22, data$BFIS_27, data$BFIS_32, 
  #                      data$BFIS_37, data$BFIS_42, data$BFIS_47, 
   #                     data$BFIS_52, data$BFIS_57)
mitems <- cbind(data$BFIS_2, data$BFIS_7, data$BFIS_12, data$BFIS_17,
                       data$BFIS_22, data$BFIS_27, data$BFIS_32, 
                        data$BFIS_37, data$BFIS_42, data$BFIS_47, 
                        data$BFIS_52, data$BFIS_57)
mkey <- c(1 , 1, -1, -1, -1, 1, 1, -1, -1, -1, 1, 1)
agree_self <- scoreItems(mkey, mitems, totals = FALSE, ilabels = NULL,missing=TRUE, impute="mean", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$agree_self <- agree_self$scores

agree_self
sd(data$agree_self)


#Extraversion
# Ext = 1, 6, 11R, 16R, 21, 26R, 31R, 36R, 41, 46, 51R, 56
mitemse <- cbind(data$BFIS_1, data$BFIS_6, data$BFIS_11, data$BFIS_16,
                 data$BFIS_21, data$BFIS_26, data$BFIS_31, data$BFIS_36, 
                 data$BFIS_41, data$BFIS_46, data$BFIS_51, 
                 data$BFIS_56)
mkeye <- c(1 , 1, -1, -1, 1, -1, -1, -1, 1, 1, -1, 1)
extra_self <- scoreItems(mkeye, mitemse, totals = FALSE, ilabels = NULL,missing=TRUE, impute="mean", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$extra_self <- extra_self$scores

extra_self
sd(data$extra_self)

#openness
# Open = 5R, 10, 15, 20, 25R, 30R, 35, 40, 45R, 50R, 55R, 60
mitemso <- cbind(data$BFIS_5, data$BFIS_10, data$BFIS_15, data$BFIS_20,
                 data$BFIS_25, data$BFIS_30, data$BFIS_35, data$BFIS_40, 
                 data$BFIS_45, data$BFIS_50, data$BFIS_55, data$BFIS_60)
mkeyo <- c(-1 , 1, 1, 1, -1, -1, 1, 1, -1, -1, -1, 1)
open_self <- scoreItems(mkeyo, mitemso, totals = FALSE, ilabels = NULL,missing=TRUE, impute="mean", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$open_self <- open_self$scores

open_self
sd(data$open_self)

#Conscientiousness
# Consc = 3R, 8R, 13, 18, 23R, 28R, 33, 38, 43, 48R, 53, 58R
mitemsc <- cbind(data$BFIS_3, data$BFIS_8, data$BFIS_13, data$BFIS_18,
                 data$BFIS_23, data$BFIS_28, data$BFIS_33, data$BFIS_38, 
                 data$BFIS_43, data$BFIS_48, data$BFIS_53, 
                 data$BFIS_58)
mkeyc <- c(-1 , -1, 1, 1, -1, -1, 1, 1, 1, -1, 1, -1)
consc_self <- scoreItems(mkeyc, mitemsc, totals = FALSE, ilabels = NULL,missing=TRUE, impute="mean", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$consc_self <- consc_self$scores

consc_self
sd(data$consc_self)

#Neuroticism
# Neur = 4R, 9R, 14, 19, 24R, 29R, 34, 39, 44R, 49R, 54, 59
mitemsn <- cbind(data$BFIS_4, data$BFIS_9, data$BFIS_14, data$BFIS_19,
                 data$BFIS_24, data$BFIS_29, data$BFIS_34, data$BFIS_39, 
                 data$BFIS_44, data$BFIS_49, data$BFIS_54, data$BFIS_59)
mkeyn <- c(1 , -1, 1, 1, -1, -1, 1, 1, -1, -1, 1, 1)
neuro_self <- scoreItems(mkeyn, mitemsn, totals = FALSE, ilabels = NULL,missing=TRUE, impute="mean", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$neuro_self<- neuro_self$scores

neuro_self
sd(data$neuro_self)

# HP = HP45r HP46r HP47r HP48r HP49r HP50 HP51 HP52
mitemshp <- cbind(data$BFIS_HP45, data$BFIS_HP46, data$BFIS_HP47,
                  data$BFIS_HP48, data$BFIS_HP49, data$BFIS_HP50,
                  data$BFIS_HP51, data$BFIS_HP52)
mkeyhp <- c(-1 , -1, -1, -1, -1, 1, 1, 1)
hp_self <- scoreItems(mkeyhp, mitemshp, totals = FALSE, ilabels = NULL, missing=TRUE, impute="mean", delete=TRUE, min = NULL, max = NULL, digits = 4)
data$hp_self <- hp_self$scores

hp_self
sd(data$hp_self)

#create dataframe with only P_ID and self-report BFI scores


#print the head of the dataframe
```

##Merge the three data frames

```{r merge}
#Merge the lab data and the 2 self-report dataframes (online, prescreen). 
#lab_data and online common variable is "ID""
#lab_data and prescreen data common variable is "P_ID" 
#I approached this by merging each of the self-report surveys with the lab data independently and then combining those dataframes. I am sure there are other (probably better) ways.

#print the head of the dataframe
```

##remove dyads with only 1 participant and repeating rows. Some participants completed the pre-measures twice. We will let R remove duplicate responses to avoid QRPs.

```{r}
#First remove any duplicate participant data

#remove any 1 person dyads

#remove multiple responses by same participant

#remove the ID column

#print the head of the dataframe
```

##Now that we have our final dataframe let's try some more advanced data manipulation.

To analyze the data in lavaan, I need to have a dataframe per trait and a row per dyad with the following columns ("dyad","p1trait", "p1percp2", "p2trait", "p2percp1") in each dataframe.

```{r}
#First create 6 data frames, one for each trait.

#Next, use dplyr functions (spread, gather, etc.) to create a row for each dyad (the participant IDs do not matter because they are exchangable). The column names should be Dyad_ID p1_self, p1_perp2, p2_self, p2_per_p1.

#print the head of the dataframe
```

##Usually we want our data in tidy (long) form. Use the dplyr functions to create a dataframe that contains all of theinformation, but only has the following columns: P_ID, Dyad_ID, rating_type (self, perception), and trait (agree, extra, open, neuro, consc). 

```{r}
#print the head of the dataframe
```



